---
title: "l03-additional-info-geom_bar"
output: html_document
---

In today's lecture there was a question about what happens when we do not use stat="Identity" in a geom_bar statement.

The code shown in lecture was using the infectious disease data in the Ch01_ID-data.csv dataset.

The plot for that looked like this:

```{r make-bar-chart}
library(readr)
id_data <- read_csv("Ch01_ID-data.csv")
library(ggplot2)
ggplot(id_data, aes(x = disease, y = percent_cases)) +
  geom_bar(stat="Identity")
```


the `stat="Identity"  designation tells the geom_bar function that we want to over-ride the default setting and use the actual values stored in `percent_cases` to determine the height of the bars.  If we do not add `stat="Identity"`  the code will not run:



You will get a message that says  "Error: stat_count() can only have an x or y aesthetic."

This is because the default for geom_bar is to take one X or Y variable and count the number of occurrences for that variable.  If we try that with our x variable, notice what happens:

```{r nostat2}
ggplot(id_data, aes(x = disease)) +
  geom_bar()
```


This is because R has "counted" the number of times each value for the variable 'disease' shows up in the dataset.  
But remember the structure of our data frame:


```{r structure}
id_data
```

Each line in the dataset captures the aggregated data for that disease, so each value for the variable `disease` appears only one time in the dataset.  

Likewise if we try to use geom_bar for `percent_cases` we get the following:


```{r nostat-cases}
ggplot(id_data, aes(percent_cases)) +
  geom_bar()
```

R is counting how many times each value for `percent_cases` appears in the data.  

If we had a dataframe/dataset where each line/row represented one reported case of an infectious disease and we had a separate column indicating the disease then we might use the geom_bar() without stat="Identity" to have R summarize for us the number of cases in each category.  

For example, if I had this dataset:

```{r }
id_data_long <- read_csv("id_data_long.csv")
head(id_data_long)
dim(id_data_long)
```
Notice the structure of the data.  Now, if I use geom_bar I would use the syntax without stat="Identity" as follows:

```{r}
ggplot(id_data_long, aes(x = disease)) +
  geom_bar()
```
