---
title: "frog-regression-example"
output: html_document
---
### Example from the text book

Read in the data on frog mating call frequency and temperature:

```{r, fig.align='center', out.width="80%"}
library(tibble)

frog_data <- tibble(id = 1:20,
       temp = c(19, 21, 22, 22, 23, 23, 23, 23, 23, 
                24, 24, 24, 24,
                25, 25, 25, 25, 26, 26, 27),
       freq = c(38, 42, 45, 45, 41, 45, 48, 50, 53, 51, 48, 53, 47,
                53, 49, 56, 53, 55, 55, 54))
```

### Scatter plot
```{r, fig.align='center', out.width="45%"}
ggplot(frog_data, aes(x = temp, y = freq)) +
  geom_point() +
  theme_minimal(base_size = 15) +
  labs(x = "Temperature (Celcius)", y = "Frequency of mating calls")
```

- Does the relationship look linear?
- Is the relationship positive or negative?

### Run the linear model
```{r, echo=TRUE}
frog_lm <- lm(formula = freq ~ temp, data = frog_data)
tidy(frog_lm)
glance(frog_lm)
```

### Check the model diagnostics

```{r}
frog_lm_augment <- augment(frog_lm)
```

- `augment()` is another `broom` function. It augments the original data frame
with the residual (`.resid`) and fitted (`.fitted`) values, among other values
that we won't talk about now.

- Make sure to know the augment command!

### Check the model diagnostics

```{r, fig.align='center', out.width="80%"}
 frog_resid<- ggplot(frog_lm_augment, aes(temp, freq)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(alpha = 0.5) +
  geom_segment(aes(xend = temp, yend = .fitted), lty = 2, alpha = 0.5) +
  theme_minimal(base_size = 15) +
  labs(title = "(a) Scatter plot")
```

### Check the model diagnostics:  residuals
```{r resid, fig.align='center', out.width="80%", echo=FALSE}
frog_resid
```

### Check the model diagnostics

```{r, fig.align='center', out.width="80%"}
# QQ plot
  frog_qq<-ggplot(frog_lm_augment, aes(sample = .resid)) + 
  geom_qq() + 
  geom_qq_line() +
  theme_minimal(base_size = 15) +
  labs(y = "Residuals", x = "Theoretical quantiles", title = "(b) QQplot")

```

### Check the model diagnostics

```{r, fig.align='center', out.width="80%", echo=FALSE}
# QQ plot
  frog_qq
```

### Check the model diagnostics

```{r, fig.align='center', out.width="80%"}
## Fitted vs. residuals
  frog_fitresid<-ggplot(frog_lm_augment, aes(y = .resid, x = .fitted)) + 
  geom_point() + 
  theme_minimal(base_size = 15) +
  geom_hline(aes(yintercept = 0)) +
  labs(y = "Residuals", x = "Fitted values", title = "(c) Fitted vs. residuals") 

```

### Check the model diagnostics

```{r, fig.align='center', out.width="80%", echo=FALSE}
## Fitted vs. residuals
  frog_fitresid
```


### Check the model diagnostics

```{r, fig.align='center', out.width="80%"}
## Amount explained
frog_gather <- frog_lm_augment %>% select(freq, .resid) %>%
  gather(key = "type", value = "value", freq, .resid)

frog_expl<-ggplot(frog_gather, aes(y = value)) + 
  geom_boxplot(aes(fill = type)) +
  theme_minimal(base_size = 15) +
  labs(title = "(d) Amount explained")
```

### Check the model diagnostics

```{r, fig.align='center', out.width="80%", echo=FALSE}
frog_expl
```

## testing and CI with linear regression models


### Part III hypothesis testing in regression models

- How can we account for the imprecision of only having a sample?
- Confidence interval for the slope 
- Hypothesis testing for the slope


### Regression standard error
All of our hypothesis tests are looking for the "signal" in the "noise".

The "noise" here is the regression standard error, the variability around our regression line.

### Regression standard error

- A good-fitting model will have a low regression standard error because $\hat{y}$
will be close to y. 
- Look at $s$ after running a linear model to assess the model's fit to the data.
- $s$ is on the same scale as y (i.e., they have the same units). 
- `glance(your_lm)` prints $s$, denoted by `sigma`

$$s = \sqrt{\frac{1}{n-2}\sum_{i=1}^{n}{\hat{e}^2}}$$
$$s = \sqrt{\frac{1}{n-2}\sum_{i=1}^{n}{(y-\hat{y})^2}}$$


### `glance()` to view the regression standard error in our frog data

```{r}
glance(frog_lm) 
```

- $sigma = 2.82$. This is the regression standard error.
- It can be interpreted as an average measure of the absolute distance between 
the observed and predicted response variable across all observations ($y$ and $\hat{y}$).

### Another way to contextualize the regression standard error

You can compute a five number summary on the residuals using the augmented data 
frame:

```{r}
frog_lm_augment %>% summarise(min_resid = min(.resid),
                            q25_resid = quantile(.resid, 0.25),
                            mean_resid = mean(.resid),
                            q75_resid = quantile(.resid, 0.75),
                            max_resid = max(.resid))
```


### Another way to contextualize the regression standard error
- The smallest residual is -6.42 and the largest is 5.58. 
- The IQR for the residuals goes from -1.92 to 1.00. 
- The mean residual is very close to 0.
- **The residual standard error (2.82) is capturing the standard deviation of this
distribution of residuals.** 
